The impact of environmental shocks on children's schooling in Bolivia

### 2020-05-08 final_seminar_paper.Rmd
finishing model specification

```{r import household data 2011, echo = F, message = F}
data_2011 <- read_dta("Raw data/seminar_paper_data/2011_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_05,
        grade = s4_06b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_06a,
        urb_rur = area,
        breakfast = s4_07,
        voucher = s4_08,
        voucher_spent = s4_09,
        educ = e,
        job = cod_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

```{r household data 2012, echo = F, message = F}
data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_04,
        grade = s4_05b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_05a,
        urb_rur = area,
        depto = departamento,
        breakfast = s4_07,
        voucher = s4_08,
        voucher_spent = s4_09,
        educ = e,
        job = cods5_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext)

@@

@@

```{r creating line graphs for variation within variables, echo = F, message = F}

Beni <- subset(flood_summary, dept == "Beni" & year > 2003, select = c("year", "floods")) 
Chuq <- subset(flood_summary, dept == "Chuquisaca" & year > 2003, select = c("year", "floods"))
Pando <- subset(flood_summary, dept == "Pando" & year > 2003, select = c("year", "floods"))
SantaCruz <- subset(flood_summary, dept == "Santa Cruz" & year > 2003, select = c("year", "floods"))
Oruro <- subset(flood_summary, dept == "Oruro" & year > 2003, select = c("year", "floods"))
Potosi <- subset(flood_summary, dept == "Potosi" & year > 2003, select = c("year", "floods"))
LaPaz <- subset(flood_summary, dept == "La Paz" & year > 2003, select = c("year", "floods"))
Tarija <- subset(flood_summary, dept == "Tarija" & year > 2003, select = c("year", "floods"))
Cocha <- subset(flood_summary, dept == "Cochabamba" & year > 2003, select = c("year", "floods"))
# Create a first line
plot(LaPaz, type = "b", frame = FALSE, pch = 19, 
     col = "red", xlab = "year", ylab = "flood counts")
# Add lines
lines(Potosi, pch = 18, col = "orange", type = "b", lty = 1)
lines(Oruro, pch = 18, col = "orangered4", type = "b", lty = 1)
lines(Beni, pch = 18, col = "green", type = "b", lty = 1)
lines(Pando, pch = 18, col = "yellow", type = "b", lty = 1)
lines(SantaCruz, pch = 18, col = "green4", type = "b", lty = 1)
lines(Cocha, pch = 18, col = "midnightblue", type = "b", lty = 1)
lines(Tarija, pch = 18, col = "mediumorchid1", type = "b", lty = 1)
lines(Chuq, pch = 18, col = "blue", type = "b", lty = 1)
# Add a legend to the plot
legend("topleft", legend=c("LaPaz", "Potosi", "Oruro", "Beni", "Pando", "SantaCruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)

@@

#Empirical Model
see the link for math equation help.

>>>>

```{r import household data 2011, echo = F, message = F}
data_2011 <- read_dta("Raw data/seminar_paper_data/2011_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_05,
        grade = s4_06b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_06a,
        urb_rur = area,
        breakfast = s4_07,
        educ = e,
        job = cod_16a,
        tasks = s5_16b,
        tasks = 
        pov_line = z,
        extr_pov = zext) %>%
  mutate(school_meal = case_when(breakfast == 1 ~ 1,
                                 breakfast == 2 ~ 0))

@@

```{r household data 2012, echo = F, message = F}
data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
  rename(hh_id = folio,
        assist1 = s4_12,  
        reason = s4_13,
        sex = s1_03,
        age = s1_04,
        relation = s1_08,
        literacy1 = s4_01,
        register1 = s4_04,
        grade = s4_05b,
        repeated = s4_11a,
        income = yhogpc,
        school_level = s4_05a,
        urb_rur = area,
        depto = departamento,
        breakfast = s4_07,
        voucher = s4_08,
        voucher_spent = s4_09,
        educ = e,
        job = cods5_16a,
        tasks = s5_16b,
        pov_line = z,
        extr_pov = zext) %>%
  mutate(school_meal = case_when(breakfast == 1 ~ 1,
                                 breakfast == 2 ~ 1,
                                 breakfast == 3 ~ 0))

@@

```{r merging 2011 and 2012 subsample for school lunch analysis}
meal_data <- data_2012 %>% 
  rbind(data_2012, data_2011)
  left_join(flood_total) %>%
  left_join(drought_total) %>%
```

@@

```{r creating line graphs for variation within flood variables, echo = F, message = F}

Beni <- subset(flood_summary, dept == "Beni" & year > 2003, select = c("year", "floods")) 
Chuq <- subset(flood_summary, dept == "Chuquisaca" & year > 2003, select = c("year", "floods"))
Pando <- subset(flood_summary, dept == "Pando" & year > 2003, select = c("year", "floods"))
SantaCruz <- subset(flood_summary, dept == "Santa Cruz" & year > 2003, select = c("year", "floods"))
Oruro <- subset(flood_summary, dept == "Oruro" & year > 2003, select = c("year", "floods"))
Potosi <- subset(flood_summary, dept == "Potosi" & year > 2003, select = c("year", "floods"))
LaPaz <- subset(flood_summary, dept == "La Paz" & year > 2003, select = c("year", "floods"))
Tarija <- subset(flood_summary, dept == "Tarija" & year > 2003, select = c("year", "floods"))
Cocha <- subset(flood_summary, dept == "Cochabamba" & year > 2003, select = c("year", "floods"))
# Create a first line
plot(LaPaz, type = "b", frame = FALSE, pch = 19, main = "Figure 2. Floods Over Time", cex.main = 1, col = "red", xlab = "year", ylab = "flood counts")
# Add lines
lines(Potosi, pch = 18, col = "orange", type = "b", lty = 1)
lines(Oruro, pch = 18, col = "orangered4", type = "b", lty = 1)
lines(Beni, pch = 18, col = "green", type = "b", lty = 1)
lines(Pando, pch = 18, col = "yellow", type = "b", lty = 1)
lines(SantaCruz, pch = 18, col = "green4", type = "b", lty = 1)
lines(Cocha, pch = 18, col = "midnightblue", type = "b", lty = 1)
lines(Tarija, pch = 18, col = "mediumorchid1", type = "b", lty = 1)
lines(Chuq, pch = 18, col = "blue", type = "b", lty = 1)
# Add a legend to the plot
legend("topleft", legend=c("La Paz", "Potosi", "Oruro", "Beni", "Pando", "Santa Cruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)

@@

```{r creating line graphs for variation within drought variables, echo = F, message = F}

Benid <- subset(drought_summary, dept == "Beni" & year > 2003, select = c("year", "droughts")) 
Chuqd <- subset(drought_summary, dept == "Chuquisaca" & year > 2003, select = c("year", "droughts"))
Pandod <- subset(drought_summary, dept == "Pando" & year > 2003, select = c("year", "droughts"))
SantaCruzd <- subset(drought_summary, dept == "Santa Cruz" & year > 2003, select = c("year", "droughts"))
Orurod <- subset(drought_summary, dept == "Oruro" & year > 2003, select = c("year", "droughts"))
Potosid <- subset(drought_summary, dept == "Potosi" & year > 2003, select = c("year", "droughts"))
LaPazd <- subset(drought_summary, dept == "La Paz" & year > 2003, select = c("year", "droughts"))
Tarijad <- subset(drought_summary, dept == "Tarija" & year > 2003, select = c("year", "droughts"))
Cochad <- subset(drought_summary, dept == "Cochabamba" & year > 2003, select = c("year", "droughts"))
# Create a first line
plot(LaPazd, type = "b", frame = FALSE, pch = 19, main = "Figure 3. Droughts Over Time", cex.main = 1, col = "red", xlab = "year", ylab = "drought counts")
# Add lines
lines(Potosid, pch = 18, col = "orange", type = "b", lty = 1)
lines(Orurod, pch = 18, col = "orangered4", type = "b", lty = 1)
lines(Benid, pch = 18, col = "green", type = "b", lty = 1)
lines(Pandod, pch = 18, col = "yellow", type = "b", lty = 1)
lines(SantaCruzd, pch = 18, col = "green4", type = "b", lty = 1)
lines(Cochad, pch = 18, col = "midnightblue", type = "b", lty = 1)
lines(Tarijad, pch = 18, col = "mediumorchid1", type = "b", lty = 1)
lines(Chuqd, pch = 18, col = "blue", type = "b", lty = 1)
# Add a legend to the plot
legend("topleft", legend=c("La Paz", "Potosi", "Oruro", "Beni", "Pando", "Santa Cruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)

#Empirical Model

The subscript 'd' refers to the department and 't' to the year, as the probit regression utilizes both department-level and year fixed-effects.

$attend_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}lny_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
